---
title: "DimorphismStatistics21"
author: "Gabrielle Bueno"
date: "4/11/2021"
output: html_document
---
# Sperm Competition

1. Load packages
```{r}
library(mosaic)
library(radiant)
library(lme4)
library(lmerTest) #Shows p-values on mixed models 
```

2. Enter the data for all testes measurements
```{r}
library(readxl)
TestesComparisons <- read_excel("~/Google Drive/Projects/Body Dimorphism/TestesComparisons.xlsx")
```

2a. Subset the data by season 
```{r}
# Separate dry from wet season
DrySeasTestes <- TestesComparisons[TestesComparisons$CaptureSeason != "2007a", ]
WetSeasTestes <- TestesComparisons[TestesComparisons$CaptureSeason == "2007a", ]
```

2b. Subset the Dry Season data by male group membership
```{r}
DrySMMales <- DrySeasTestes[DrySeasTestes$GroupComp == "SM", ]
DryMMMales <- DrySeasTestes[DrySeasTestes$GroupComp == "MM", ]
DryMMDom <- DryMMMales[DryMMMales$ChestStain == "Stained", ]
DryMMSubsCI <- DryMMMales[DryMMMales$ChestStain != "Stained", ]
DryMMClean <- DryMMMales[DryMMMales$ChestStain == "Clean", ]
DryMMInter <- DryMMMales[DryMMMales$ChestStain == "Intermediate", ]
DryAllDom <- DrySeasTestes[DrySeasTestes$ChestStain == "Stained", ]
```

2c. Subset data by presence of groupmates (paired sets)
```{r}
PairedAllSeas <- TestesComparisons[TestesComparisons$GroupMates == "Y", ]
PairedDry <- DrySeasTestes[DrySeasTestes$GroupMates == "Y", ]
```

## 3. Rank: Total Testes Mass

3a. Total Testes Mass by Rank in Multi-Male Groups: LMER
```{r}
lmmodMMTotal <- lmer(TotalTestesMass ~ Rank + (1|LemurID) + (1|Group), data = DryMMMales)
summary(lmmodMMTotal)
```

3b: Total Testes Mass by Rank in Multi-Male Groups: Descriptive Statistics
```{r}
group_by(DryMMMales, Rank) %>%
  summarise(
    median = median(TotalTestesMass, na.rm = TRUE),
    mean = mean(TotalTestesMass, na.rm = TRUE),
    IQR = IQR(TotalTestesMass, na.rm = TRUE),
    sd = sd(TotalTestesMass, na.rm = TRUE)
  )
```

## 4.  Rank: Relative Testes Mass

4. Relative Testes Mass by Rank in Multi-Male Groups: LMER
```{r}
lmmodMMRelative <- lmer(RelativeTestesMass ~ Rank + (1|LemurID) + (1|Group), data = DryMMMales)
summary(lmmodMMRelative)
```

## 5. Rank: Weight

5. Weight by Rank in Multi-Male Groups: LMER
```{r}
lmmodMMWeights <- lmer(Weight ~ Rank + (1|LemurID) + (1|Group), data = DryMMMales)
summary(lmmodMMWeights)
```

5b. Weight by Rank in Multi-Male Groups: Descriptive Statistics
```{r}
group_by(DryMMMales, Rank) %>%
  summarise(
    median = median(Weight, na.rm = TRUE),
    mean = mean(Weight, na.rm = TRUE),
    IQR = IQR(Weight, na.rm = TRUE),
    sd = sd(Weight, na.rm = TRUE)
  )
```

## 6. Group Comp: Total Testes Mass

6. Total Testes Mass by Group Composition (SM vs MM Dominants): LMER
```{r}
lmmodDomTotal <- lmer(TotalTestesMass ~ GroupComp + (1|LemurID), data = DryAllDom)
summary(lmmodDomTotal)
```

## 7. Group Comp: Relative Testes Mass

7. Relative Testes Mass by Group Composition (SM vs MM Dominants): LMER
```{r}
lmmodDomRelative <- lmer(RelativeTestesMass ~ GroupComp + (1|LemurID), data = DryAllDom)
summary(lmmodDomRelative)
```
## 8. Group Comp: Weight

8. Weight by Group Composition (SM vs MM Dominants): LMER
```{r}
lmmodDomWeights <- lmer(Weight ~ GroupComp + (1|LemurID), data = DryAllDom)
summary(lmmodDomWeights)
```

8b. Weight by Group Composition (SM vs MM Dominants): Descriptive Statistics
```{r}
group_by(DryAllDom, GroupComp) %>%
  summarise(
    median = median(Weight, na.rm = TRUE),
    mean = mean(Weight, na.rm = TRUE),
    IQR = IQR(Weight, na.rm = TRUE),
    sd = sd(Weight, na.rm = TRUE)
  )
```

## 9. Averages Across Groups

9. Looking at averages across groups: descriptive stats
```{r}
dryG2 <- DryMMMales[DryMMMales$Group == "II", ]
dryG3 <- DryMMMales[DryMMMales$Group == "III", ]
dryG4 <- DryMMMales[DryMMMales$Group == "IV", ]
dryGUnmarked <- DryMMMales[DryMMMales$Group == "Unmarked", ] #only 1 male
dryG6 <- DryMMMales[DryMMMales$Group == "VI", ] # only 1 male
```
9a. Group 2 (n = 7, nDom = 4, nSub = 3)
```{r}
group_by(dryG2, Rank) %>%
  summarise(
    median = median(TotalTestesMass, na.rm = TRUE),
    mean = mean(TotalTestesMass, na.rm = TRUE),
    IQR = IQR(TotalTestesMass, na.rm = TRUE),
    sd = sd(TotalTestesMass, na.rm = TRUE)
  )
```
9b. Group 3 (n = 7, nDom = 3, nSub = 4)
```{r}
group_by(dryG3, Rank) %>%
  summarise(
    median = median(TotalTestesMass, na.rm = TRUE),
    mean = mean(TotalTestesMass, na.rm = TRUE),
    IQR = IQR(TotalTestesMass, na.rm = TRUE),
    sd = sd(TotalTestesMass, na.rm = TRUE)
  )
```
9c. Group 4 (n = 6, nDom = 3, nSub = 3)
```{r}
group_by(dryG4, Rank) %>%
  summarise(
    median = median(TotalTestesMass, na.rm = TRUE),
    mean = mean(TotalTestesMass, na.rm = TRUE),
    IQR = IQR(TotalTestesMass, na.rm = TRUE),
    sd = sd(TotalTestesMass, na.rm = TRUE)
  )
```
9d. Group 6 (n = 1, nSub = 1)
```{r}
# Only Orion, Subordinate/Intermediate
group_by(dryG6, Rank) %>%
  summarise(
    mean = mean(TotalTestesMass, na.rm = TRUE)
  )
```
9e. Unmarked Group (n = 1, nSub = 1)
```{r}
# Only Isaac, Subordinate
group_by(dryGUnmarked, Rank) %>%
  summarise(
    mean = mean(TotalTestesMass, na.rm = TRUE)
  )
```


# Fecundity

Packages
```{r}
library(dplyr)
library(lme4)
library(lmerTest)
library(glmmML)
library(nlme)
```

1. Import Data
```{r}
library(readxl)
AllAdultFemales <- read_excel("~/Google Drive/Projects/Fecundity/AllAdultFemales.xlsx")
```

2. Subset Data to Just Adult Females 
```{r}
dummy <- AllAdultFemales[AllAdultFemales$CaptureSeason != "2007a", ]
dummy$AvgTailCrown <- as.numeric(dummy$AvgTailCrown)
AdultFemales <- dummy[dummy$Age >= 5, ]
AdultFemalesIncl4 <- dummy[dummy$Age >= 4, ]
belowAdult <- dummy[dummy$Age <= 4, ]
```

## 3. Did they give birth? 

3a. WEIGHT: Did they give birth? Including Age 4 (which is technically 5)

```{r}
#AdultFemalesIncl4$GaveBirth <- factor(AdultFemalesIncl4$GaveBirth)
AdultFemalesIncl4$LemurID <- factor(AdultFemalesIncl4$LemurID)
lmAFBirthWeight <- glmer(GaveBirthNum ~ Weight + (1|LemurID), data = AdultFemalesIncl4, family = binomial)
summary(lmAFBirthWeight)
```

3a. Visualize
```{r}
group_by(AdultFemalesIncl4, GaveBirth) %>%
  summarise(
    median = median(Weight, na.rm = TRUE),
    mean = mean(Weight, na.rm = TRUE),
    IQR = IQR(Weight, na.rm = TRUE),
    sd = sd(Weight, na.rm = TRUE)
  )
```

3b. TAIL-CROWN: Did they give birth? Including Age 4
No significance 
```{r}
lmAFBirthTC <- lmer(AvgTailCrown ~ GaveBirth + (1|LemurID), data = AdultFemalesIncl4)
summary(lmAFBirthTC)
```

3b. Visualize
```{r}
group_by(AdultFemalesIncl4, GaveBirth) %>%
  summarise(
    median = median(AvgTailCrown, na.rm = TRUE),
    mean = mean(AvgTailCrown, na.rm = TRUE),
    IQR = IQR(AvgTailCrown, na.rm = TRUE),
    sd = sd(AvgTailCrown, na.rm = TRUE)
  )
```

3c. WEIGHT: Did they give birth? Only 5+
Significant but less so
```{r}
lmAF5BirthWeight <- lmer(Weight ~ GaveBirth + (1|LemurID), data = AdultFemales)
summary(lmAF5BirthWeight)
```

3c. Visualize
```{r}
group_by(AdultFemales, GaveBirth) %>%
  summarise(
    median = median(Weight, na.rm = TRUE),
    mean = mean(Weight, na.rm = TRUE),
    IQR = IQR(Weight, na.rm = TRUE),
    sd = sd(Weight, na.rm = TRUE)
  )
```

## 4. Infant Survival to Age 1

4a. WEIGHT: Infant Survival to Age 1: Including Age 4
Not Significant
```{r}
lmAFInfSurv <- lmer(Weight ~ SurvivedOneYear + (1|LemurID), data = AdultFemalesIncl4)
summary(lmAFInfSurv)
```

4a. Visualize
```{r}
group_by(AdultFemalesIncl4, SurvivedOneYear) %>%
  summarise(
    median = median(Weight, na.rm = TRUE),
    mean = mean(Weight, na.rm = TRUE),
    IQR = IQR(Weight, na.rm = TRUE),
    sd = sd(Weight, na.rm = TRUE)
  )
```

4b. TAIL-CROWN: Infant Survival to Age 1: Including Age 4
Not Significant
```{r}
lmAFInfSurvTC <- lmer(AvgTailCrown ~ SurvivedOneYear + (1|LemurID), data = AdultFemalesIncl4)
summary(lmAFInfSurvTC)
```

4b. Visualize
```{r}
group_by(AdultFemales, SurvivedOneYear) %>%
  summarise(
    median = median(AvgTailCrown, na.rm = TRUE),
    mean = mean(AvgTailCrown, na.rm = TRUE),
    IQR = IQR(AvgTailCrown, na.rm = TRUE),
    sd = sd(AvgTailCrown, na.rm = TRUE)
  )
```

4c. WEIGHT: Infant Survival to Age 1: Only 5+
Not Significant
```{r}
lmAFInfSurv5 <- lmer(Weight ~ SurvivedOneYear + (1|LemurID), data = AdultFemales)
summary(lmAFInfSurv5)
```

4c. Visualize
```{r}
group_by(AdultFemales, SurvivedOneYear) %>%
  summarise(
    median = median(Weight, na.rm = TRUE),
    mean = mean(Weight, na.rm = TRUE),
    IQR = IQR(Weight, na.rm = TRUE),
    sd = sd(Weight, na.rm = TRUE)
  )
```

## 5. Infant Weight at Age 1ish

5a. WEIGHT: Infant Weight at Age 1ish: Including Age 4
```{r}
AdultFemalesIncl4$InfWeightAge <- as.numeric(AdultFemalesIncl4$InfWeightAge)
lmAFInfWt <- lmer(InfWeightAge ~ Weight + (1|LemurID), data = filter(AdultFemalesIncl4, !is.na(InfWeightAge)))
summary(lmAFInfWt)
```


5b. TAIL-CROWN: Infant Weight at Age 1ish: Including Age 4

5c. WEIGHT: Infant Weight at Age 1ish: Only 5+