---
title: "GeneralDimorphismCode"
output: html_document
---

# Weight and Nested ANOVAs 

Boxplot
```
library(ggplot2)
ggplot(BodyMassData, aes(x=Age, y=Weight, fill=Sex)) + geom_boxplot()
```
Nested ANOVA

```
library(car)
###Convert the independent variables to factors for the ANOVA to group by
BodyMassData$Sex<-as.factor(BodyMassData$Sex)
BodyMassData$Age<-as.factor(BodyMassData$Age)
```

Look at boxplot to check the assumption of normality
```
boxplot(Weight~Sex,data=BodyMassData)
boxplot(Weight~Age,data=BodyMassData)
```

Fitting the model

* Sex is nested within age, for the example site is nested within season
* Use aov function to fit an ANOVA
* To indicate that sex is nested within age we use a / operation in the aov call:
```
model.BodyMassData<-aov(Weight~Age/Sex,data=BodyMassData)
####OR####
model.BodyMassData<-aov(Weight~Age+Sex%in%Age,data=BodyMassData)
```

to look at the results, call the summary function on the model object I just created
```
summary(model.BodyMassData)
##Summary results:
##              Df Sum Sq Mean Sq F value Pr(>F)    
##Age           5  33.50   6.700 105.336 <2e-16 ***
##Age:Sex       6   0.36   0.059   0.932  0.475    
##Residuals   108   6.87   0.064                   
```

Explanation of the results

* 2 null hypotheses in an ANOVA. First is that there's no effect of A, second is that there's no effect of B nested within A.
* P-values show there's a significant result for Age, but not for Sex nested in Age.
* So can reject the null hypothesis that there's no effect of Age, but fail to reject the second null hypothesis.


**DIAGNOSTIC: evaluate the residuals**

Make a residuals vs fitted values plot
```
plot(model.BodyMassData$residuals~model.BodyMassData$fitted.values)
```

I want to do a t-test for just ages 5, so I've uploaded a new excel file 
```
attach(AdultsOnly)
```

Just looking
```
boxplot(Weight~Sex)
```

* H0: mean weight of females = mean weight of males
* two sided test
* assume non-equal variances

* mu=0 -> means null hypothesis is that the mean difference is 0
* alt -> alternative being used is a two-sided
* confidence interval of 95%
* assuming variances are equal and that these two groups aren't paired
```
t.test(Weight~Sex, mu=0, alt="two.sided", conf=0.95, var.eq=F, paired=F, data=AdultsOnly)
```
Decide on variance in two ways:
1. Look at the boxplot, see if they have similar variations in weight (F seems more)
2. Compare the actual variances--see below
``` 
var(Weight[Sex=="F"])
var(Weight[Sex=="M"])
```

* Calculations from end of sperm competition, to examined female and male weight. 
Adding females
```{r}
library(readxl)
WeightComparisons326 <- read_excel("~/Google Drive/Projects/Body Dimorphism/WeightComparisons326.xlsx")
```

Subsetting
```{r}
FemAndMM <- WeightComparisons326[with(WeightComparisons326, Sex == "F" | GroupComp == "MM"), ]
FemAndSM <- WeightComparisons326[with(WeightComparisons326, Sex == "F" | GroupComp == "SM"), ]
```

Females and MM
```{r}
group_by(FemAndMM, Sex) %>%
  summarise(
    median = median(Weight, na.rm = TRUE),
    mean = mean(Weight, na.rm = TRUE),
    IQR = IQR(Weight, na.rm = TRUE),
    SD = sd(Weight, na.rm = TRUE)
  )
```
```{r}
wilcox.test(Weight ~ Sex, data = FemAndMM)
```





# Looking at weight of all males--line chart
```
ggplot(AdultMales, aes(x=CaptureDate, y=Weight, color=LemurID)) +geom_line()
```



# Leg Length

Data Notes

* Micka was excluded--his leg length is too variable

1. I've uploaded `LegLengthAdultAvg`. 

2. Let's subset it into male and female leg length averages
```
FLeg <- LegLengthAdultAvg[LegLengthAdultAvg$Sex == "F", ]
MLeg <- LegLengthAdultAvg[LegLengthAdultAvg$Sex == "M", ]
```
3. First, I'm running a Shapiro test to test normality (because I will be using a one-tailed independent t-test, so I must test normality and variance)
```
shapiro.test(MLeg$AvgLegLength)

	Shapiro-Wilk normality test

data:  MLeg$AvgLegLength
W = 0.96085, p-value = 0.3256
#######
shapiro.test(FLeg$AvgLegLength)

	Shapiro-Wilk normality test

data:  FLeg$AvgLegLength
W = 0.89398, p-value = 0.06441
```
Both p-values are greater than 0.05, so I can assume normality.

4. Now I'm running an f-test to test the variance
```
var.test(AvgLegLength ~ Sex, data = LegLengthAdultAvg)
############
	F test to compare two variances

data:  AvgLegLength by Sex
F = 1.557, num df = 15, denom df = 29, p-value = 0.2982
alternative hypothesis: true ratio of variances is not equal to 1
95 percent confidence interval:
 0.6697303 4.1286781
sample estimates:
ratio of variances 
             1.557 
```
The p-value is greater than 0.05, so there is no significant difference between the variances. 




# Sperm Competition: Actually Working on (Aaron look here!)

This section is for more focused models of the sperm competition data. The longer one below has more in it (including wet season data, which I will not be looking at because the sample sizes are so small) and is messier. The organization of this one is as follows: 
1. Data entry
2. Subsetting data
3. By rank in MM groups, all males: Absolute Testes Mass
4. By rank in MM groups, all males: Relative Testes Mass
5. By group composition, all males: Absolute Testes Mass
6. By group composition, all males: Relative Testes Mass
7. By rank in MM groups, averages: Absolute Testes Mass
8. By rank in MM groups, averages: Relative Testes Mass
9. By group composition, averages: Absolute Testes Mass
10. By group composition, averages: Relative Testes Mass

Add section for loading libraries, only use CSVs, change names from dry/wet to mating/birthing


Data Notes
* Micka 2008 is intermediate in a SM group: he was transitioning, he was stained by August. He's the NA
* Omby 2007a is excluded until we can get pictures 
* 2006 Quincy has been edited to be clean
* 2018 Zoma is excluded, as he's an adult natal male 

1a. Load packages
```{r}
library(mosaic)
library(radiant)
library(lme4)
library(lmerTest) #Shows p-values on mixed models 
```
1b. Enter the data for all testes measurements
```{r}
library(readxl)
TestesComparisons <- read_excel("~/Google Drive/Projects/Body Dimorphism/TestesComparisons.xlsx")
```

1c. Enter the data for testes averages
* Note: For this, I average the testes size (and weight) for each individual male when he was measured more than once in a group/dominance class, so that I could run t-tests. So for example Quincy is measured twice as a subordinate male in a multi-male group, and I averaged his relative and total testes mass for those two times. I have also excluded all wet season data from this dataset. This data is subset in 5b
```{r}
library(readxl)
TestesAverages <- read_excel("~/Google Drive/Projects/Body Dimorphism/TestesAverages.xlsx")
```

2a. Subset the data by season **#change name to breeding vs wet season**
```{r}
# Separate dry from wet season
DrySeasTestes <- TestesComparisons[TestesComparisons$CaptureSeason != "2007a", ]
WetSeasTestes <- TestesComparisons[TestesComparisons$CaptureSeason == "2007a", ]
```

2b. Subset the Dry Season data by male group membership
```{r}
DrySMMales <- DrySeasTestes[DrySeasTestes$GroupComp == "SM", ]
DryMMMales <- DrySeasTestes[DrySeasTestes$GroupComp == "MM", ]
DryMMDom <- DryMMMales[DryMMMales$ChestStain == "Stained", ]
DryMMSubsCI <- DryMMMales[DryMMMales$ChestStain != "Stained", ]
DryMMClean <- DryMMMales[DryMMMales$ChestStain == "Clean", ]
DryMMInter <- DryMMMales[DryMMMales$ChestStain == "Intermediate", ]
DryAllDom <- DrySeasTestes[DrySeasTestes$ChestStain == "Stained", ]
```

2c. Subset data by presence of groupmates (paired sets)
```{r}
PairedAllSeas <- TestesComparisons[TestesComparisons$GroupMates == "Y", ]
PairedDry <- DrySeasTestes[DrySeasTestes$GroupMates == "Y", ]
```

2d. Subset averaged data by male group membership and dominance
```{r}
AveragedAllDom <- TestesAverages[TestesAverages$ChestStain == "Stained", ]
AveragedMMMales <- TestesAverages[TestesAverages$GroupComp == "MM", ]
AveragedSMMales <- TestesAverages[TestesAverages$GroupComp == "SM", ]
```

## 3. By rank in MM groups, all males: Absolute Testes Mass

3a. Total Testes Mass by Rank in MM groups (Dry) -- Mann-Whitney test
```{r}
group_by(DryMMMales, Rank) %>%
  summarise(
    median = median(TotalTestesMass, na.rm = TRUE),
    mean = mean(TotalTestesMass, na.rm = TRUE),
    IQR = IQR(TotalTestesMass, na.rm = TRUE),
    sd = sd(TotalTestesMass, na.rm = TRUE)
  )
```
```{r}
wilcox.test(TotalTestesMass ~ Rank, data = DryMMMales)
```

3b. Plot for Total Testes Size by Rank in MM groups
```{r}
ggplot(data = DryMMMales, aes(x = Rank, y = TotalTestesMass)) + geom_point()
```

3c. Regression like a t-test -- same results as a t-test
```{r}
modMMTotal <- lm(TotalTestesMass ~ Rank, data = DryMMMales)
#modMMTotal #RankSub is the slope 
summary(modMMTotal)
```

3d. Linear mixed model like a t-test + repeated measures
```{r}
library(lme4)
```
```{r}
lmmodMMTotal <- lmer(TotalTestesMass ~ Rank + (1|LemurID), data = DryMMMales)
summary(lmmodMMTotal)
```


## 4. By rank in MM groups, all males: Relative Testes Mass

4a. Relative Testes Mass by Rank in MM groups (Dry): Mann-Whitney Test
```{r}
group_by(DryMMMales, Rank) %>%
  summarise(
    median = median(RelativeTestesMass, na.rm = TRUE),
    mean = mean(RelativeTestesMass, na.rm = TRUE),
    IQR = IQR(RelativeTestesMass, na.rm = TRUE),
    sd = sd(RelativeTestesMass, na.rm = TRUE)
  )
```
```{r}
wilcox.test(RelativeTestesMass ~ Rank, data = DryMMMales)
```

4b. Plot for Relative Testes Size by Rank in MM groups
```{r}
ggplot(data = DryMMMales, aes(x = Rank, y = RelativeTestesMass)) + geom_point()
```

4c. Linear mixed model like a t-test + repeated measures
```{r}
lmmodMMRelative <- lmer(RelativeTestesMass ~ Rank + (1|LemurID), data = DryMMMales)
summary(lmmodMMRelative)
```

4d. Weight by Rank in MM groups (Dry): Mann-Whitney
A: Not significant
```{r}
group_by(DryMMMales, Rank) %>%
  summarise(
    median = median(Weight, na.rm = TRUE),
    mean = mean(Weight, na.rm = TRUE),
    IQR = IQR(Weight, na.rm = TRUE),
    sd = sd(Weight, na.rm = TRUE)
  )
```
```{r}
wilcox.test(Weight ~ Rank, data = DryMMMales)
```

4e. Weight by Rank in MM Groups (Dry): lmer
```{r}
lmmodMMWeights <- lmer(Weight ~ Rank + (1|LemurID), data = DryMMMales)
summary(lmmodMMWeights)
```

## 5. By group composition, all males: Absolute Testes Mass

5a. Absolute Testes Mass Comparing SM Dominants and MM Dominants (Dry): Mann-Whitney and T-Test
```{r}
group_by(DryAllDom, GroupComp) %>%
  summarise(
    median = median(TotalTestesMass, na.rm = TRUE),
    mean = mean(TotalTestesMass, na.rm = TRUE),
    IQR = IQR(TotalTestesMass, na.rm = TRUE),
    sd = sd(TotalTestesMass, na.rm = TRUE)
  )
```
```{r}
wilcox.test(TotalTestesMass ~ GroupComp, data = DryAllDom)
t.test(TotalTestesMass~GroupComp, data = DryAllDom)
```

5b. Plot for Absolute Testes Size by Group Composition, Dominant Males
```{r}
ggplot(data = DryAllDom, aes(x = GroupComp, y = TotalTestesMass)) + geom_point()
```

5c. Linear mixed effects model like a t-test + repeated measures
```{r}
lmmodDomsTotal <- lmer(TotalTestesMass ~ GroupComp + (1|LemurID), data = DryAllDom)
summary(lmmodDomsTotal)
```

##Look
5d. LMM including body mass 
A: This means that when I control for weight, dom males in single male groups have smaller testes by 0.43
```{r}
allDomWeightFixedMod <- lmer(TotalTestesMass ~ GroupComp + Weight + (1|LemurID), data = DryAllDom)
summary(allDomWeightFixedMod)
```


## 6. By group composition, all males: Relative Testes Mass

6a. Relative Testes Mass Comparing SM Dominants and MM Dominants (Dry): Mann-Whitney and T-Test
```{r}
group_by(DryAllDom, GroupComp) %>%
  summarise(
    median = median(RelativeTestesMass, na.rm = TRUE),
    mean = mean(RelativeTestesMass, na.rm = TRUE),
    IQR = IQR(RelativeTestesMass, na.rm = TRUE),
    sd = sd(RelativeTestesMass, na.rm = TRUE)
  )
```
```{r}
wilcox.test(RelativeTestesMass ~ GroupComp, data = DryAllDom)
t.test(RelativeTestesMass~GroupComp, data = DryAllDom)
```

6b. Plot for Relative Testes Size by Group Composition, Dominant Males
```{r}
ggplot(data = DryAllDom, aes(x = GroupComp, y = RelativeTestesMass)) + geom_point()
```

6c. Linear mixed effects model like a t-test + repeated measures
```{r}
lmmodDomsRelative <- lmer(RelativeTestesMass ~ GroupComp + (1|LemurID), data = DryAllDom)
summary(lmmodDomsRelative)
```

6d. Weight by Group Composition, Dominant Males (Dry): Mann-Whitney
A: Not significant
```{r}
group_by(DryAllDom, GroupComp) %>%
  summarise(
    median = median(Weight, na.rm = TRUE),
    mean = mean(Weight, na.rm = TRUE),
    IQR = IQR(Weight, na.rm = TRUE),
    sd = sd(Weight, na.rm = TRUE)
  )
```
```{r}
wilcox.test(Weight ~ GroupComp, data = DryAllDom)
```

6e. Weight by Group Composition, Dominant Males: lmer
```{r}
lmmodDomsWeights <- lmer(Weight ~ GroupComp + (1|LemurID), data = DryAllDom)
summary(lmmodDomsWeights)
```

## 7. By rank in MM groups, averages: Absolute Testes Mass

7a. Looking at Male Averages: Total Testes Mass by Rank in MM groups: t-test
```{r}
group_by(AveragedMMMales, Rank) %>%
  summarise(
    median = median(AvgTotalTestes, na.rm = TRUE),
    mean = mean(AvgTotalTestes, na.rm = TRUE),
    IQR = IQR(AvgTotalTestes, na.rm = TRUE)
  )
```
Checking normality: all normal!
```{r}
shapiro.test(AveragedMMMales$AvgTotalTestes)
var.test(AvgTotalTestes ~ Rank, data = AveragedMMMales) #F-test
```
```{r}
t.test(AvgTotalTestes ~ Rank, data = AveragedMMMales)
```

## 8. By rank in MM groups, averages: Relative Testes Mass

8a. Looking at Male Averages: Relative Testes Mass by Rank in MM groups
```{r}
group_by(AveragedMMMales, Rank) %>%
  summarise(
    median = median(AvgRelativeTestes, na.rm = TRUE),
    mean = mean(AvgRelativeTestes, na.rm = TRUE),
    IQR = IQR(AvgRelativeTestes, na.rm = TRUE)
  )
```
Checking normality: all normal!
```{r}
shapiro.test(AveragedMMMales$AvgRelativeTestes)
var.test(AvgRelativeTestes ~ Rank, data = AveragedMMMales) #F-test
```
```{r}
t.test(AvgRelativeTestes ~ Rank, data = AveragedMMMales)
```

## 9. By group composition, averages: Absolute Testes Mass

9a. Looking at Male Averages: Total Testes Mass by Group Composition, Just Dominants
```{r}
group_by(AveragedAllDom, GroupComp) %>%
  summarise(
    median = median(AvgTotalTestes, na.rm = TRUE),
    mean = mean(AvgTotalTestes, na.rm = TRUE),
    IQR = IQR(AvgTotalTestes, na.rm = TRUE)
  )
```
Checking normality: all normal!
```{r}
shapiro.test(AveragedAllDom$AvgTotalTestes)
var.test(AvgTotalTestes ~ GroupComp, data = AveragedAllDom) #F-test
```
```{r}
t.test(AvgTotalTestes ~ GroupComp, data = AveragedAllDom)
```


## 10. By group composition, averages: Relative Testes Mass

10a. Looking at Male Averages: Relative Testes Mass by Group Composition, Just Dominants
```{r}
group_by(AveragedAllDom, GroupComp) %>%
  summarise(
    median = median(AvgRelativeTestes, na.rm = TRUE),
    mean = mean(AvgRelativeTestes, na.rm = TRUE),
    IQR = IQR(AvgRelativeTestes, na.rm = TRUE)
  )
```
Checking normality: all normal!
```{r}
shapiro.test(AveragedAllDom$AvgRelativeTestes)
var.test(AvgRelativeTestes ~ GroupComp, data = AveragedAllDom) #F-test
```
```{r}
t.test(AvgRelativeTestes ~ GroupComp, data = AveragedAllDom)
```


## 11. Other things we worked on

11a. Linear mixed model with multi-male and single-male data
```{r}
lmmAllDryAllRanks <- lmer(TotalTestesMass ~ Rank + GroupComp + (1|LemurID), data=DrySeasTestes)
summary(lmmAllDryAllRanks)
```

11b. Linear mixed model with multi-male and single-male data, with an interaction by rank
```{r}
lmmAllDryRankInter <- lmer(TotalTestesMass ~ Rank * GroupComp + (1|LemurID), data=DrySeasTestes)
summary(lmmAllDryRankInter)
```


11c. Plot for paired males
```{r}
ggplot(data = PairedDry, aes(x = Rank, y = TotalTestesMass, color = CaptureSeason, group = CaptureSeason)) + geom_point() + geom_line()
```

12a. Testes size predicted by body mass -- All, absolute
```{r, fig.show="hold", out.width="50%"}
testesByMassAllPlot1 <- ggplot(data = DrySeasTestes, aes(x = Weight, y = TotalTestesMass, color = LemurID)) + geom_point()
#testesByMassAllPlot1

testesbyMassAllPlot2 <- ggplot(data = DryMMMales, aes(x = Weight, y = TotalTestesMass, color = LemurID)) + geom_point()
testesbyMassAllPlot2
testesbyMassAllPlot3 <- ggplot(data = DrySMMales, aes(x = Weight, y = TotalTestesMass, color = LemurID)) + geom_point()
testesbyMassAllPlot3
```

12b. Testes size by body mass -- averages


13. Testes size by age absolute
```{r}
ggplot(data = DrySeasTestes, aes(x = Age, y = RelativeTestesMass, color = LemurID)) + geom_point()
```



***



# Sperm Competition: Everything

Organization Notes
* Numbering is a bit off, it was too much work to go back and renumber each time I added something. I'll fix it at the end. 

Data Notes
* Micka 2008 is intermediate in a SM group: he was transitioning, he was stained by August. He's the NA
* Omby 2007a is excluded until we can get pictures 
* 2006 Quincy has been edited to be clean
* 2018 Zoma is excluded, as he's an adult natal male 

1. Enter the data for all testes measurements
```{r}
library(readxl)
TestesComparisons <- read_excel("~/Google Drive/Projects/Body Dimorphism/TestesComparisons.xlsx")
```

1b. Enter the data for testes averages
* Note: For this, I average the testes size for each individual male when he was measured more than once in a group/dominance class, so that I could run t-tests. So for example Quincy is measured twice as a subordinate male in a multi-male group, and I averaged his relative and total testes mass for those two times. I have also excluded all wet season data from this dataset. This data is subset in 5b
```{r}
library(readxl)
TestesAverages <- read_excel("~/Google Drive/Projects/Body Dimorphism/TestesAverages.xlsx")
```
2. Subset the data by season
```{r}
# Separate dry from wet season
DrySeasTestes <- TestesComparisons[TestesComparisons$CaptureSeason != "2007a", ]
WetSeasTestes <- TestesComparisons[TestesComparisons$CaptureSeason == "2007a", ]
```

3. Subset the Dry Season data by male group membership
```{r}
DrySMMales <- DrySeasTestes[DrySeasTestes$GroupComp == "SM", ]
DryMMMales <- DrySeasTestes[DrySeasTestes$GroupComp == "MM", ]
DryMMDom <- DryMMMales[DryMMMales$ChestStain == "Stained", ]
DryMMSubsCI <- DryMMMales[DryMMMales$ChestStain != "Stained", ]
DryMMClean <- DryMMMales[DryMMMales$ChestStain == "Clean", ]
DryMMInter <- DryMMMales[DryMMMales$ChestStain == "Intermediate", ]
DryAllDom <- DrySeasTestes[DrySeasTestes$ChestStain == "Stained", ]
```

4. Subset the Wet Season (January 2007) data
```{r}
JanSM <- WetSeasTestes[WetSeasTestes$GroupComp == "SM", ]
JanMM <- WetSeasTestes[WetSeasTestes$GroupComp == "MM", ]
JanMMDom <- JanMM[JanMM$ChestStain == "Stained", ]
JanMMSub <- JanMM[JanMM$ChestStain == "Clean", ]
JanAllDom <- WetSeasTestes[WetSeasTestes$ChestStain == "Stained", ]
```

5. Subset all data by group membership, in case I want to include January 2007
```{r}
AllSeasSM <- TestesComparisons[TestesComparisons$GroupComp == "SM", ]
AllSeasMM <- TestesComparisons[TestesComparisons$GroupComp == "MM", ]
AllSeasMMDom <- AllSeasMM[AllSeasMM$ChestStain == "Stained", ]
AllSeasMMSubsCI <- AllSeasMM[AllSeasMM$ChestStain != "Stained", ]
AllSeasMMClean <- AllSeasMM[AllSeasMM$ChestStain == "Clean", ]
AllSeasMMInter <- AllSeasMM[AllSeasMM$ChestStain == "Intermediate", ]
AllSeasAllDom <- TestesComparisons[TestesComparisons$ChestStain == "Stained", ]
```

5a. Subset data by presence of groupmates (paired sets)
```{r}
PairedAllSeas <- TestesComparisons[TestesComparisons$GroupMates == "Y", ]
PairedDry <- DrySeasTestes[DrySeasTestes$GroupMates == "Y", ]
PairedWet <- WetSeasTestes[WetSeasTestes$GroupMates == "Y", ]
```

5b. Subset averaged data by male group membership and dominance
```{r}
AveragedAllDom <- TestesAverages[TestesAverages$ChestStain == "Stained", ]
AveragedMMMales <- TestesAverages[TestesAverages$GroupComp == "MM", ]
AveragedSMMales <- TestesAverages[TestesAverages$GroupComp == "SM", ]
```

5c. Plot for paired males
```{r}
ggplot(data = PairedDry, aes(x = Rank, y = TotalTestesMass, color = CaptureSeason, group = CaptureSeason)) + geom_point() + geom_line()
```

6. Total Testes Mass by Rank in MM groups (Dry)
```{r}
group_by(DryMMMales, Rank) %>%
  summarise(
    median = median(TotalTestesMass, na.rm = TRUE),
    mean = mean(TotalTestesMass, na.rm = TRUE),
    IQR = IQR(TotalTestesMass, na.rm = TRUE),
    sd = sd(TotalTestesMass, na.rm = TRUE)
  )
```
```{r}
wilcox.test(TotalTestesMass ~ Rank, data = DryMMMales)
```

6b. Plot for Total Testes Size by Rank in MM groups
```{r}
ggplot(data = DryMMMales, aes(x = Rank, y = TotalTestesMass)) + geom_point()
```

6c. Regression like a t-test -- same results as a t-test
```{r}
modMMTotal <- lm(TotalTestesMass ~ Rank, data = DryMMMales)
#modMMTotal #RankSub is the slope 
summary(modMMTotal)
```

6d. Linear mixed model like a t-test + repeated measures
```{r}
library(lme4)
```
```{r}
lmmodMMTotal <- lmer(TotalTestesMass ~ Rank + (1|LemurID), data = DryMMMales)
summary(lmmodMMTotal)
```

6e. Linear mixed model with multi-male and single-male data
```{r}
lmmAllDryAllRanks <- lmer(TotalTestesMass ~ Rank + GroupComp + (1|LemurID), data=DrySeasTestes)
summary(lmmAllDryAllRanks)
```

6f. Linear mixed model with multi-male and single-male data, with an interaction by rank
```{r}
lmmAllDryRankInter <- lmer(TotalTestesMass ~ Rank * GroupComp + (1|LemurID), data=DrySeasTestes)
summary(lmmAllDryRankInter)
```

7. Relative Testes Mass by Rank in MM groups (Dry)
```{r}
group_by(DryMMMales, Rank) %>%
  summarise(
    median = median(RelativeTestesMass, na.rm = TRUE),
    mean = mean(RelativeTestesMass, na.rm = TRUE),
    IQR = IQR(RelativeTestesMass, na.rm = TRUE),
    sd = sd(RelativeTestesMass, na.rm = TRUE)
  )
```
```{r}
wilcox.test(RelativeTestesMass ~ Rank, data = DryMMMales)
```

7b. Weight by Rank in MM groups (Dry)
A: Not significant
```{r}
group_by(DryMMMales, Rank) %>%
  summarise(
    median = median(Weight, na.rm = TRUE),
    mean = mean(Weight, na.rm = TRUE),
    IQR = IQR(Weight, na.rm = TRUE),
    sd = sd(Weight, na.rm = TRUE)
  )
```
```{r}
wilcox.test(Weight ~ Rank, data = DryMMMales)
```

8. Total Testes Mass for MM groups including January 2007 data
```{r}
group_by(AllSeasMM, Rank) %>%
  summarise(
    median = median(TotalTestesMass, na.rm = TRUE),
    mean = mean(TotalTestesMass, na.rm = TRUE),
    IQR = IQR(TotalTestesMass, na.rm = TRUE)
  )
```
```{r}
wilcox.test(TotalTestesMass ~ Rank, data = AllSeasMM)
```
9. Relative Testes Mass for MM groups including January 2007 data
```{r}
group_by(AllSeasMM, Rank) %>%
  summarise(
    median = median(RelativeTestesMass, na.rm = TRUE),
    mean = mean(RelativeTestesMass, na.rm = TRUE),
    IQR = IQR(RelativeTestesMass, na.rm = TRUE)
  )
```
```{r}
wilcox.test(RelativeTestesMass ~ Rank, data = AllSeasMM)
```

10. Absolute Testes Mass Comparing SM Dominants and MM Dominants (Dry)
```{r}
group_by(DryAllDom, GroupComp) %>%
  summarise(
    median = median(TotalTestesMass, na.rm = TRUE),
    mean = mean(TotalTestesMass, na.rm = TRUE),
    IQR = IQR(TotalTestesMass, na.rm = TRUE),
    sd = sd(TotalTestesMass, na.rm = TRUE)
  )
```
```{r}
wilcox.test(TotalTestesMass ~ GroupComp, data = DryAllDom)
t.test(TotalTestesMass~GroupComp, data = DryAllDom)
```

11. Relative Testes Mass Comparing SM Dominants and MM Dominants (Dry)
```{r}
group_by(DryAllDom, GroupComp) %>%
  summarise(
    median = median(RelativeTestesMass, na.rm = TRUE),
    mean = mean(RelativeTestesMass, na.rm = TRUE),
    IQR = IQR(RelativeTestesMass, na.rm = TRUE)
  )
```
```{r}
wilcox.test(RelativeTestesMass ~ GroupComp, data = DryAllDom)
```
11a. Weight Comparing SM Dominants and MM Dominants (Dry)
```{r}
group_by(DryAllDom, GroupComp) %>%
  summarise(
    median = median(Weight, na.rm = TRUE),
    mean = mean(Weight, na.rm = TRUE),
    IQR = IQR(Weight, na.rm = TRUE),
    sd = sd(Weight, na.rm = TRUE)
  )
```
```{r}
wilcox.test(Weight ~ GroupComp, data = DryAllDom)
```

12. Total Testes Mass Comparing SM Dominants and MM Dominants (All)
```{r}
group_by(AllSeasAllDom, GroupComp) %>%
  summarise(
    median = median(TotalTestesMass, na.rm = TRUE),
    mean = mean(TotalTestesMass, na.rm = TRUE),
    IQR = IQR(TotalTestesMass, na.rm = TRUE)
  )
```
```{r}
wilcox.test(TotalTestesMass ~ GroupComp, data = AllSeasAllDom)
```

13. Relative Testes Mass Comparing SM Dominants and MM Dominants (All)
```{r}
group_by(AllSeasAllDom, GroupComp) %>%
  summarise(
    median = median(RelativeTestesMass, na.rm = TRUE),
    mean = mean(RelativeTestesMass, na.rm = TRUE),
    IQR = IQR(RelativeTestesMass, na.rm = TRUE)
  )
```
```{r}
wilcox.test(RelativeTestesMass ~ GroupComp, data = AllSeasAllDom)
```

14. Total Testes Mass Comparing Doms and Subs in Paired Groups (Dry Seasons)
A: T-test is significant, wilcox isn't
```{r}
group_by(PairedDry, Rank) %>%
  summarise(
    median = median(TotalTestesMass, na.rm = TRUE),
    mean = mean(TotalTestesMass, na.rm = TRUE),
    IQR = IQR(TotalTestesMass, na.rm = TRUE)
  )
```
Checking normality: all normal!
```{r}
shapiro.test(PairedDry$TotalTestesMass)
var.test(TotalTestesMass ~ Rank, data = PairedDry) #F-test
```
```{r}
wilcox.test(TotalTestesMass ~ Rank, data = PairedDry)
t.test(TotalTestesMass ~ Rank, data = PairedDry, var.equal = TRUE)
```
##Look
14c. Paired graph by indiv, absolute
```{r}
ggplot(data = PairedDry, aes(x = Rank, y = TotalTestesMass, color = LemurID, group = LemurID)) + geom_point() + geom_line()
```
14d. Paired graph by indiv, relative
```{r}
ggplot(data = PairedDry, aes(x = Rank, y = RelativeTestesMass, color = LemurID, group = LemurID)) + geom_point() + geom_line()
```

15. Relative Testes Mass Comparing Doms and Subs in Paired Groups (Dry Seasons)
A: Not significant (with Zoma, Wilcox is significant, but t-test isn't)
```{r}
group_by(PairedDry, Rank) %>%
  summarise(
    median = median(RelativeTestesMass, na.rm = TRUE),
    mean = mean(RelativeTestesMass, na.rm = TRUE),
    IQR = IQR(RelativeTestesMass, na.rm = TRUE)
  )
```
Checking normality: all normal!
```{r}
shapiro.test(PairedDry$RelativeTestesMass)
var.test(RelativeTestesMass ~ Rank, data = PairedDry) #F-test
```
```{r}
wilcox.test(RelativeTestesMass ~ Rank, data = PairedDry)
t.test(RelativeTestesMass ~ Rank, data = PairedDry, var.equal = TRUE)
```
16. All MM males compared to all SM males: Absolute Testes Mass
A: Not significant
```{r}
group_by(DrySeasTestes, GroupComp) %>%
  summarise(
    median = median(TotalTestesMass, na.rm = TRUE),
    mean = mean(TotalTestesMass, na.rm = TRUE),
    IQR = IQR(TotalTestesMass, na.rm = TRUE),
    SD = sd(TotalTestesMass, na.rm = TRUE)
  )
```
```{r}
wilcox.test(TotalTestesMass ~ GroupComp, data = DrySeasTestes)
```

17. All MM males compared to all SM males: Relative Testes Mass
A: Not significant
```{r}
group_by(DrySeasTestes, GroupComp) %>%
  summarise(
    median = median(RelativeTestesMass, na.rm = TRUE),
    mean = mean(RelativeTestesMass, na.rm = TRUE),
    IQR = IQR(RelativeTestesMass, na.rm = TRUE),
    SD = sd(RelativeTestesMass, na.rm = TRUE)
  )
```
```{r}
wilcox.test(RelativeTestesMass ~ GroupComp, data = DrySeasTestes)
```

18. All MM males compared to all SM males: Weight
```{r}
group_by(DrySeasTestes, GroupComp) %>%
  summarise(
    median = median(Weight, na.rm = TRUE),
    mean = mean(Weight, na.rm = TRUE),
    IQR = IQR(Weight, na.rm = TRUE),
    SD = sd(Weight, na.rm = TRUE)
  )
```
```{r}
wilcox.test(Weight ~ GroupComp, data = DrySeasTestes)
```

19. Looking at Male Averages: Total Testes Mass by Rank in MM groups
```{r}
group_by(AveragedMMMales, Rank) %>%
  summarise(
    median = median(AvgTotalTestes, na.rm = TRUE),
    mean = mean(AvgTotalTestes, na.rm = TRUE),
    IQR = IQR(AvgTotalTestes, na.rm = TRUE)
  )
```
Checking normality: all normal!
```{r}
shapiro.test(AveragedMMMales$AvgTotalTestes)
var.test(AvgTotalTestes ~ Rank, data = AveragedMMMales) #F-test
```
```{r}
t.test(AvgTotalTestes ~ Rank, data = AveragedMMMales)
```
19a. Looking at Male Averages: Relative Testes Mass by Rank in MM groups
```{r}
group_by(AveragedMMMales, Rank) %>%
  summarise(
    median = median(AvgRelativeTestes, na.rm = TRUE),
    mean = mean(AvgRelativeTestes, na.rm = TRUE),
    IQR = IQR(AvgRelativeTestes, na.rm = TRUE)
  )
```
Checking normality: all normal!
```{r}
shapiro.test(AveragedMMMales$AvgRelativeTestes)
var.test(AvgRelativeTestes ~ Rank, data = AveragedMMMales) #F-test
```
```{r}
t.test(AvgRelativeTestes ~ Rank, data = AveragedMMMales)
```
19c. Looking at Male Averages: Total Testes Mass by Group Composition, Just Dominants
```{r}
group_by(AveragedAllDom, GroupComp) %>%
  summarise(
    median = median(AvgTotalTestes, na.rm = TRUE),
    mean = mean(AvgTotalTestes, na.rm = TRUE),
    IQR = IQR(AvgTotalTestes, na.rm = TRUE)
  )
```
Checking normality: all normal!
```{r}
shapiro.test(AveragedAllDom$AvgTotalTestes)
var.test(AvgTotalTestes ~ GroupComp, data = AveragedAllDom) #F-test
```
```{r}
t.test(AvgTotalTestes ~ GroupComp, data = AveragedAllDom)
```

19d. Looking at Male Averages: Relative Testes Mass by Group Composition, Just Dominants
```{r}
group_by(AveragedAllDom, GroupComp) %>%
  summarise(
    median = median(AvgRelativeTestes, na.rm = TRUE),
    mean = mean(AvgRelativeTestes, na.rm = TRUE),
    IQR = IQR(AvgRelativeTestes, na.rm = TRUE)
  )
```
Checking normality: all normal!
```{r}
shapiro.test(AveragedAllDom$AvgRelativeTestes)
var.test(AvgRelativeTestes ~ GroupComp, data = AveragedAllDom) #F-test
```
```{r}
t.test(AvgRelativeTestes ~ GroupComp, data = AveragedAllDom)
```










***
Adding females
```{r}
library(readxl)
WeightComparisons326 <- read_excel("~/Google Drive/Projects/Body Dimorphism/WeightComparisons326.xlsx")
```

Subsetting
```{r}
FemAndMM <- WeightComparisons326[with(WeightComparisons326, Sex == "F" | GroupComp == "MM"), ]
FemAndSM <- WeightComparisons326[with(WeightComparisons326, Sex == "F" | GroupComp == "SM"), ]
```

Females and MM
```{r}
group_by(FemAndMM, Sex) %>%
  summarise(
    median = median(Weight, na.rm = TRUE),
    mean = mean(Weight, na.rm = TRUE),
    IQR = IQR(Weight, na.rm = TRUE),
    SD = sd(Weight, na.rm = TRUE)
  )
```
```{r}
wilcox.test(Weight ~ Sex, data = FemAndMM)
```


