---
title: "StatsQuestions"
output: html_document
---



# Sperm Competition

1. Is the W of a wilcoxon rank sum exact test the same as the U of a Mann-Whitney? 

1. Load packages
```{r}
library(mosaic)
library(lme4)
library(lmerTest)
library(readxl)
```
2. Read in data (sorry, tried to do this with git but couldn't get it to work)
```{r}
TestesDrySeason <- read_excel("~/Google Drive/Projects/Body Dimorphism/TestesDrySeason.xlsx")
TestesAverages <- read_excel("~/Google Drive/Projects/Body Dimorphism/TestesAverages.xlsx")
```

3a. Subset Dry Season Data
```{r}
DrySMMales <- TestesDrySeason[TestesDrySeason$GroupComp == "SM", ]
DryMMMales <- TestesDrySeason[TestesDrySeason$GroupComp == "MM", ]
DryAllDom <- TestesDrySeason[TestesDrySeason$ChestStain == "Stained", ]
```

3b. Subset Averaged Data
```{r}
AveragedAllDom <- TestesAverages[TestesAverages$ChestStain == "Stained", ]
AveragedMMMales <- TestesAverages[TestesAverages$GroupComp == "MM", ]
AveragedSMMales <- TestesAverages[TestesAverages$GroupComp == "SM", ]
```


4. Total Testes Mass by Rank in Multi-Male Groups
4a. Visualize
```{r}
ggplot(data = DryMMMales, aes(x = Rank, y = TotalTestesMass)) + geom_point()
```

4b. Linear Mixed Effects Model
```{r}
lmmodMMTotal <- lmer(TotalTestesMass ~ Rank + (1|LemurID) + (1|Group), data = DryMMMales)
summary(lmmodMMTotal)
```

```{r}
lmmodMMGMTotal <- lmer(TotalTestesMass ~ Rank + (1|LemurID) + (1|GroupMateID), data = DryMMMales)
summary(lmmodMMTotal)
```



5. Total Testes Mass Comparing SM Dominants and MM Dominants

5a. Linear Mixed Effects Model
```{r}
lmmodDomsTotal <- lmer(TotalTestesMass ~ GroupComp + (1|LemurID), data = DryAllDom)
summary(lmmodDomsTotal)
```


# Female Fecundity

1. Import Data
```{r}
library(readxl)
AllAdultFemales <- read_excel("~/Google Drive/Projects/Fecundity/AllAdultFemales.xlsx")
```

2. Subset Data to Just Adult Females 
```{r}
dummy <- AllAdultFemales[AllAdultFemales$CaptureSeason != "2007a", ]
AdultFemales <- dummy[dummy$Age >= 5, ]
belowAdult <- dummy[dummy$Age <= 4, ]
```

LMER try
```{r}
lmAFPregnant <- lmer(Weight ~ Pregnant + (1|LemurID) + (1|Age), data = AdultFemales)
summary(lmAFPregnant)
```

Visualize
```{r}
group_by(AdultFemales, Pregnant) %>%
  summarise(
    median = median(Weight, na.rm = TRUE),
    mean = mean(Weight, na.rm = TRUE),
    IQR = IQR(Weight, na.rm = TRUE),
    sd = sd(Weight, na.rm = TRUE)
  )
```

```{r}
tryYoung <- lm(Weight ~ Pregnant, data = belowAdult)
summary(tryYoung)
```



Looking at tail-crown
```{r}
ggplot(data=AdultFemales, aes(x = CaptureSeason, y = TailCrown, color = LemurID, group = LemurID)) + geom_point() + geom_line()

```




# Leg Length

Trying to figure out the best t-test. I ran the following and I'm not sure which is best. 


2. This is what I used previously. In the past I put var.eq=F, but I tested the variance and they are equal, so which should I do? It seems like the second is Welch's but the first isn't? 
```
> t.test(AvgLegLength~Sex, mu=0, alt="two.sided", conf=0.95, var.eq=T, paired=F, data=LegLengthAdultAvg)

	Two Sample t-test

data:  AvgLegLength by Sex
t = 4.4452, df = 44, p-value = 5.875e-05
alternative hypothesis: true difference in means is not equal to 0
95 percent confidence interval:
 0.9512152 2.5291419
sample estimates:
mean in group F mean in group M 
       33.92462        32.18444 

> t.test(AvgLegLength~Sex, mu=0, alt="two.sided", conf=0.95, var.eq=F, paired=F, data=LegLengthAdultAvg)

	Welch Two Sample t-test

data:  AvgLegLength by Sex
t = 4.1529, df = 25.489, p-value = 0.0003237
alternative hypothesis: true difference in means is not equal to 0
95 percent confidence interval:
 0.8780208 2.6023363
sample estimates:
mean in group F mean in group M 
       33.92462        32.18444 
```

